Appendix

### 2023-05-24 manuscript/_appendix/_model-details.qmd
Add new corrections to discussion and reorganise project structure

```{r apafg}
#| label: apafg-prior-mean
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 5
#| apa-cap: "Expected prior-predicted mean."
nd <- expand.grid(n_phon_std = 0,
                  age_std = (10:36-mean(responses$age))/(sd(responses$age)),
                  exposure_std = 0,
                  lv_std = 0)

@@

prior_epreds |> 
    ggplot(aes(x = age_std, y = .epred)) +
    facet_wrap(~.category) +
    geom_line(linewidth = 3/4,
              alpha = 1/2,
              aes(group = .draw),
              colour = "grey50") + 
    scale_x_continuous(breaks = scale(seq(10, 36, 4), 
                                      mean(responses$age), 
                                      sd(responses$age))[, 1],
                       labels = seq(10, 36, 4)) +
    scale_fill_manual(values = rev(clrs), 
                      labels = c("95%", "89%", "67%", "50%")) +
    labs(x = "Age (months)", 
         y = "Probability of acquisition",
         colour = "Credible interval", 
         fill = "Credible interval") +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(legend.position = "top",
          axis.ticks.x = element_line(colour = "black"),
          panel.grid = element_blank())
```

>>>>

```{r apafg}
#| label: fig-prior-mean
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 3
#| out-width: 90%
#| fig-cap: "Expected prior-predicted mean."
nd <- expand.grid(n_phon_std = 0,
                  age_std = (10:36-mean(responses$age))/(sd(responses$age)),
                  exposure_std = 0,
                  lv_std = 0)

@@

prior_epreds |> 
    ggplot(aes(x = age_std, y = .epred)) +
    facet_wrap(~ .category) +
    geom_line(linewidth = 3/4,
              alpha = 1/2,
              aes(group = .draw),
              colour = "grey50") + 
    labs(x = "Age (months)", 
         y = "Probability of acquisition",
         colour = "Credible interval", 
         fill = "Credible interval") +
    scale_x_continuous(breaks = scale(seq(10, 36, 4), 
                                      mean(responses$age), 
                                      sd(responses$age))[, 1],
                       labels = seq(10, 36, 4)) +
    theme(legend.position = "top",
          axis.ticks.x = element_line(colour = "black"),
          panel.grid = element_blank())
```

### 2023-05-24 manuscript/_appendix/_model-doe.qmd
Add new corrections to discussion and reorganise project structure

```{r fig-coefs-doe}
#| label: fig-coefs-doe
#| fig-cap: "Posterior distribution of regression coefficients in the probability scale"
#| fig-note: "Median: median of the posterior distribution in the logit scale. 95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| echo: false
#| message: false
#| warning: false
#| out-width: 100%
#| fig-height: 3
#| fig-width: 12
# summarise posterior draws
bind_rows("Exposure model" = posterior_draws$summary,
          "DoE-Frequency model" = posterior_draws_doe,
          .id = "model") |> 
    mutate(
        across(c(.median, .lower, .upper),
               function(x) {
                   ifelse(
                       .type=="Intercepts (at 22 months)",
                       plogis(x),
                       x/4
                   )
               }
        )) |>
    filter(.type == "Slopes") |> 
    ggplot(aes(.median, .variable_name,
               xmin = .lower, xmax = .upper)) +
    facet_wrap(~model, scales = "free") +
    annotate(geom = "rect",
             ymin = -Inf,
             ymax = Inf,
             xmin = -0.1/4,
             xmax = +0.1/4,
             fill = "grey70") +
    geom_errorbar(linewidth = 1, width = 0.25) +
    geom_point(size = 2.5) +
    labs(x = "Median and 95% HDI of the posterior distribution",
         y = "Predictor",
         colour = "Model") +
    scale_x_continuous(labels = percent) +
    scale_colour_manual(values = clrs[c(1, 4)]) +
    theme(legend.position = "none",
          axis.title.y = element_blank(),
          panel.grid.major.y = element_blank())
```

>>>>

```{r fig-coefs-doe}
#| label: fig-coefs-doe
#| fig-cap: "Posterior distribution of regression coefficients in the probability scale"
#| fig-note: "Median: median of the posterior distribution in the logit scale. 95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| echo: false
#| message: false
#| warning: false
#| out-width: 100%
#| fig-height: 4
#| fig-width: 7
# summarise posterior draws
bind_rows("Exposure model" = model_summary,
          "DoE-Frequency model" = posterior_doe_summary,
          .id = "model") |> 
    mutate(
        across(c(.median, .lower, .upper),
               function(x) {
                   ifelse(
                       .type=="Intercepts (at 22 months)",
                       plogis(x),
                       x/4
                   )
               }
        )) |>
    filter(.type == "Slopes") |> 
    ggplot(aes(.median, .variable_name,
               xmin = .lower, xmax = .upper,
               colour = model,
               shape = model)) +
    annotate(geom = "rect",
             ymin = -Inf,
             ymax = Inf,
             xmin = -0.1/4,
             xmax = +0.1/4,
             fill = "grey90") +
    geom_vline(xintercept = 0, 
               linetype = "dashed",
               linewidth = 3/4) +
    geom_errorbar(linewidth = 3/4, 
                  position = position_dodge(width = 0.5),
                  width = 0.2) +
    geom_point(size = 2,
               position = position_dodge(width = 0.5)) +
    labs(x = "Median and 95% HDI of the posterior distribution",
         y = "Predictor",
         colour = "Model") +
    scale_colour_grey(start = 0, end = 0.6) +
    theme(legend.position = "none",
          axis.title.y = element_blank(),
          panel.grid.major.y = element_blank())
```

### 2023-05-24 manuscript/_appendix/_model-predictors.qmd
Add new corrections to discussion and reorganise project structure

```{r sample-sizes}
#| label: sample-sizes
#| echo: false
#| message: false
#| warning: false
#| include: false
tar_load(c(responses, bvq_data, ))
n_obs <- nrow(responses)

@@

## Word exposure index  {.unnumbered}

Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database [@macwhinney2000childes; @sanchez2019childes]. Using the corresponding lexical frequencies from the Catalan and Spanish corpora was not possible due to the low number of Catalan participants and tokens available in those languages. Available words in the English corpora were mapped to their Catalan and Spanish translations [see @fourtassi2020growth for a similar approach], and transformed to Zipf scores [@vanheuven2014subtlexuk; @zipf1949human]. Responses to words with missing lexical frequencies were excluded from analyses of the total number of items. Participant degree of exposure to the words' language were calculated as the percentage of exposure to the language, as reported by caretakers in the language profile section of the questionnaire. For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%.

@@

## Levenshtein similarity {.unnumbered}

Generally, a pair of words from two different languages that share meaning---translation equivalents---are considered as *cognates* if they share etymological origin (e.g., *table* and *taula*, in English and Catalan). The fact that two words share etymological origin frequently leads to them also being form-similar, as reflected by their overlapping orthographic or phonological form. This makes many cognates perceptually similar. In psycholinguistics, the term *cognateness* is often used to refer to the form-similarity that a pair of translation equivalents share, which has been found to impact how bilinguals process such word-forms (e.g., @costa2000cognate; @spivey1999cross). Whether two form-similar translation equivalents are etymologically related or not is arguably tangential to the question of how cognateness affect language processing. For instance, the translation pair *much* and *mucho*, in English and Spanish have the same meaning, and are considerably similar at the orthographical and phonological level. Nonetheless, they come from different Proto-Indo-European roots [@campbell2007glossary]. This does not keep bilinguals from processing such kind of word-forms word forms differently than other translation equivalents with no form-similarity like 'dog' and *perro*, in English and Spanish. In the scope of this experiment, we consider any form-similar translation equivalent as a *cognate* pair, regardless of whether both word-forms share etymological origin.

@@

A way to achieve this is to use phonological transcriptions of the word forms. Phonological transcriptions are symbolic, visual representations of the sounds in speech that capture some minimal  set of its features following some standard. The International Phonetic Alphabet (IPA) is one instance of such standards, providing a set of symbols that correspond to specific phonemes, as defined by their position across a series of dimensions. Using IPA transcriptions of word forms, one can use the same similarity/distance metric as in the case of orthographic word-forms. This method has been successfully used to measure the pairwise similarity between phonological word-forms from the  same language [e.g., @fourtassi2020growth] and across languages [@heeringa2004measuring; @floccia2018methods], and also between orthographic word-forms [@schepens2012distributions].

Multiple algorithms have been created to compute the distance between two strings of characters. The Levenshtein distance being one of them. The this algorithm calculates the minimum number of edit operations (additions, deletions, and substitutions) that one string must go through to become identical to the other string [@levenshtein1966binary] (see @eq-levenshtein). For instance, /\textipa{"ka.za}/ (*house*, in Catalan), would need to go through one substitution to become identical to its translation equivalent in Spanish /\textipa{"ka.sa}/ (/z/ is replaced by /s/). Therefore, the Levenshtein distance between /\textipa{"ka.za}/ and /\textipa{"ka.sa}/ is one. For an easier handling of non-ASCII characters (fairly prevalent in Catalan and Spanish, e.g., á, ü, ñ), we used phonological transcriptions in X-SAMPA format, as opposed to IPA, for its computer-friendly set of symbols [@wells1995computer].

@@

To measure the phonological similarity between the translation equivalents in our study, we computed a normalised versions of the Levenshtein distance between each pair of word-forms that accounts for the lenght of the strings. The rationale behind this correction is that longer word-forms are more likely to differ from their counterpart, compared to shorter word forms, and that this tendency does not correspond necessarily to a larger distance in terms of perception. This normalisation is achieved by dividing the Levenshtein distance by the length of the longest string, which leads to a distance metric that ranges from 0 to 1. This can be interpreted as a proportion of characters in the longest string that need to be edited in order for that string to become identical to the shorter string. A 0% normalised Levenshtein distance indicates that both word-forms are identical. A 50% normalised Levenshtein distance indicates that half of the phonemes in the longest word-form must be edited for both word-forms to become identical. A 100% normalised Levenshtein distance indicates that the two word-forms are completely different, as *all* phonemes in the longest word-form must be changed for it to become identical to the other.

@@

The R package `stringdist` [@van-der-loo2014stringdist] offers the `stringsim()` function to compute this normalised Levenshtein similarity measure, see @tbl-translation-levenshtein for more examples:

```{r tbl-translation-levenshtein}
#| label: tbl-translation-levenshtein
#| tbl-cap: Normalised Levenshtein similarity computed for three exemplars of translation pairs in our study
tribble(~Catalan, ~Spanish, ~Levenshtein,
        "porta [pOrt5]", "puerta [pwe4ta]", 0.50,
        "taula [tawl5]", "mesa [mesa]", 0.00,
        "cotxe [kOtS5]", "coche [kotSe]", 0.40) |> 
    gt() |> 
    fmt_number(is.numeric)
```

>>>>

```{r sample-sizes}
#| label: sample-sizes
#| echo: false
#| message: false
#| warning: false
#| include: false
tar_load(c(responses, bvq_data))
n_obs <- nrow(responses)

@@

### Word exposure index  {.unnumbered}

Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database [@macwhinney2000childes; @sanchez2019childesdb]. Using the corresponding lexical frequencies from the Catalan and Spanish corpora was not possible due to the low number of Catalan participants and tokens available in those languages. Available words in the English corpora were mapped to their Catalan and Spanish translations [see @fourtassi2020growth for a similar approach], and transformed to Zipf scores [@vanheuven2014subtlexuk; @zipf1949human]. Responses to words with missing lexical frequencies were excluded from analyses of the total number of items. Participant degree of exposure to the words' language were calculated as the percentage of exposure to the language, as reported by caretakers in the language profile section of the questionnaire. For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%.

@@

### Levenshtein similarity {.unnumbered}

Generally, a pair of words from two different languages that share meaning---translation equivalents---are considered as *cognates* if they share etymological origin (e.g., *table* and *taula*, in English and Catalan). The fact that two words share etymological origin frequently leads to them also being form-similar, as reflected by their overlapping orthographic or phonological form. This makes many cognates perceptually similar. In psycholinguistics, the term *cognateness* is often used to refer to the form-similarity that a pair of translation equivalents share, which has been found to impact how bilinguals process such word-forms [e.g., @costa2000cognate; @spivey1999cross]. Whether two form-similar translation equivalents are etymologically related or not is arguably tangential to the question of how cognateness affect language processing. For instance, the translation pair *much* and *mucho*, in English and Spanish have the same meaning, and are considerably similar at the orthographical and phonological level. Nonetheless, they come from different Proto-Indo-European roots [@campbell2007glossary]. This does not keep bilinguals from processing such kind of word-forms word forms differently than other translation equivalents with no form-similarity like 'dog' and *perro*, in English and Spanish. In the scope of this experiment, we consider any form-similar translation equivalent as a *cognate* pair, regardless of whether both word-forms share etymological origin.

@@

A way to achieve this is to use phonological transcriptions of the word forms. Phonological transcriptions are symbolic, visual representations of the sounds in speech that capture some minimal  set of its features following some standard. The International Phonetic Alphabet (IPA) is one instance of such standards, providing a set of symbols that correspond to specific phonemes, as defined by their position across a series of dimensions. Using IPA transcriptions of word forms, one can use the same similarity/distance metric as in the case of orthographic word-forms. This method has been successfully used to measure the pairwise similarity between phonological word-forms from the  same language [e.g., @fourtassi2020growth] and across languages [@heeringa2004measuring; @floccia2018ii], and also between orthographic word-forms [@schepens2012distributions].

Multiple algorithms have been created to compute the distance between two strings of characters. The Levenshtein distance being one of them. The this algorithm calculates the minimum number of edit operations (additions, deletions, and substitutions) that one string must go through to become identical to the other string [@levenshtein1966binary] (see @eq-levenshtein). For instance, /\textipa{"ka.za}/ (*house*, in Catalan), would need to go through one substitution to become identical to its translation equivalent in Spanish /\textipa{"ka.sa}/ (/z/ is replaced by /s/). Therefore, the Levenshtein distance between /\textipa{"ka.za}/ and /\textipa{"ka.sa}/ is one. For an easier handling of non-ASCII characters (fairly prevalent in Catalan and Spanish, e.g., á, ü, ñ), we used phonological transcriptions in X-SAMPA format, as opposed to IPA, for its computer-friendly set of symbols [@wells1995computercoding].

@@

To measure the phonological similarity between the translation equivalents in our study, we computed a normalised versions of the Levenshtein distance between each pair of word-forms that accounts for the length of the strings. The rationale behind this correction is that longer word-forms are more likely to differ from their counterpart, compared to shorter word forms, and that this tendency does not correspond necessarily to a larger distance in terms of perception. This normalisation is achieved by dividing the Levenshtein distance by the length of the longest string, which leads to a distance metric that ranges from 0 to 1. This can be interpreted as a proportion of characters in the longest string that need to be edited in order for that string to become identical to the shorter string. A 0% normalised Levenshtein distance indicates that both word-forms are identical. A 50% normalised Levenshtein distance indicates that half of the phonemes in the longest word-form must be edited for both word-forms to become identical. A 100% normalised Levenshtein distance indicates that the two word-forms are completely different, as *all* phonemes in the longest word-form must be changed for it to become identical to the other.

@@

The R package `stringdist` [@vanderloo2014stringdist] offers the `stringsim()` function to compute this normalised Levenshtein similarity measure, see @tbl-translation-levenshtein for more examples:

```{r tbl-translation-levenshtein}
#| label: tbl-translation-levenshtein
#| tbl-cap: Normalised Levenshtein similarity computed for three exemplars of translation pairs in our study
#| echo: false
tribble(~Catalan, ~Spanish, ~Levenshtein,
        "porta /pOrt5/", "puerta /pwe4ta/", 0.50,
        "taula /tawl5/", "mesa /mesa/", 0.00,
        "cotxe /kOtS5/", "coche /kotSe/", 0.40) |> 
    gt() |> 
    fmt_number(is.numeric)
```

### 2023-05-24 manuscript/_appendix/_syllable-frequency.qmd
Add new corrections to discussion and reorganise project structure

items <- items |> 
    mutate(xsampa = ipa(ipa)) |> 
    mutate(syll = strsplit(xsampa, 
                           split = paste0(.sep = c("\\.", "\\\""), 
                                          collapse = "|")),
           syll = lapply(syll, function(x) x[x != ""]),
           n_syll = map_int(syll, length),
           item = str_remove(item, "cat_|spa_"))
```

@@

```{r tbl-items-table}
#| label: tbl-syll-items
#| tbl-cap: Sample of items included in the BVQ questionnaire and their syllabified SAMPA transcriptions in Catalan and Spanish
#| echo: false
#| warning: false
#| message: false
items |> 
    distinct(te, item, language, meaning, xsampa, n_syll) |> 
    left_join(distinct(pool_tmp, te, language, item, label),
              by = join_by(te, item, language)) |> 
    pivot_wider(names_from = language,
                values_from = c(item, label, xsampa, n_syll),
                id_cols = c(te, meaning),
                names_repair = make_clean_names,
                values_fn = first) |> 
    sample_n(15) |> 
    gt() |> 
    cols_hide(c(te, item_catalan, item_spanish)) |> 
    tab_spanner("Catalan", matches("catalan")) |> 
    tab_spanner("Spanish", matches("spanish")) |> 
    cols_label(meaning = "Translation",
               label_catalan = "Item",
               xsampa_catalan = "X-SAMPA",
               n_syll_catalan = "Syllables",
               label_spanish = "Item",
               xsampa_spanish = "X-SAMPA",
               n_syll_spanish = "Syllables") |> 
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners()) |> 
    tab_style(cell_text(style = "italic"),
              cells_column_labels())
```

@@

```{r fig-syll-number}
#| label: fig-syll-number
#| fig-cap: Distribution of the number of syllables in Catalan and Spanish
#| fig-width: 7
#| fig-height: 3
#| echo: false
#| warning: false
#| message: false
pool_tmp |> 
    group_by(language) |> 
    add_count(language, name = "n_lang") |> 
    ungroup() |> 
    count(language, n_lang, n_syll) |> 
    mutate(prop = n/n_lang) |> 
    ggplot(aes(n_syll, n)) +
    facet_wrap(~ language) +
    geom_col(colour = "white", fill = "grey70") +
    geom_text(aes(label = percent(prop)),
              size = 3,
              position = position_nudge(y = 15),
              colour = "black") +
    labs(x = "Number of syllables",
         y = "Number of words",
         fill = "Language") +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())
```

@@

```{r fig-top-syll-freq}
#| label: fig-top-syll-freq
#| fig-cap: "Top 20 most frequent syllables"
#| fig-width: 8
#| fig-height: 3
#| out-width: 80%
#| echo: false
#| warning: false
#| message: false
syll |> 
    group_by(language) |> 
    top_n(20, freq_syll) |> 
    ggplot(aes(reorder_within(syll, desc(freq_syll), language),
               freq_syll)) +
    facet_wrap(~ language,
               scales = "free_x",
               ncol = 1) +
    geom_col(colour = "white",
             fill = "grey70") +
    labs(x = "Syllable",
         y = "Relative frequency\n(counts per million)",
         colour = "Language",
         fill = "Language") +
    scale_x_reordered() +
    scale_y_continuous(labels = \(x) format(x, big.mark = ",")) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 10),
          panel.grid = element_blank())
```

@@

```{r syllables-posterior}
#| label: syllables-posterior
#| echo: false
#| warning: false
#| message: false
rope_interval <- c(lower = -0.5, upper = 0.5)
# tidy predictor names
str_repl <- c(
    "b_Intercept" = "Intercept",
    "b_n_syll_std" = glue("Syllables (+1 SD, {round(sd(syllables_data$n_syll), 3)})"),
    "b_lv_std" = glue("Cognateness (+1 SD, {percent(sd(syllables_data$lv), accuracy = 0.01)})")
)

post_syllables <- model_fit_syllables |>
    describe_posterior(test = "rope",
                       rope_range = rope_interval,
                       ci_method = "HDI") |>
    as_tibble() |>
    clean_names() |>
    mutate(.variable_name = factor(
        parameter,
        levels = names(str_repl),
        labels = str_repl) |>
            as.character(),
        parameter = ifelse(str_detect(parameter, "Intercept"),
                           str_remove_all(parameter, "Intercept \\(|\\)"),
                           parameter)) |>
    select(.variable = parameter,
           .variable_name,
           .median = median,
           .upper = ci_high,
           .lower = ci_low,
           .rope_overlap = rope_percentage) |>
    ungroup()

@@

@tbl-syll-marginal shows the median posterior-predicted syllabic frequencies for words with one to four syllables, for the whole range of cognateness values. Overall, cognate words' syllabic frequency is equivalent to that of non-cognates. This suggests that the cognate facilitation effect in word acquisition reported in the present study is not the result from an association between cognateness and higher syllabic frequencies.

```{r tbl-syll-marginal}
#| label: tbl-syll-marginal
#| tbl-cap: "Posterior-predictions of the syllabic frequency model. Thicker lines indicate the median of the posterior predictions, and thinner lines indicate individual posterior predictions."
#| fig-width: 7
#| fig-height: 5
#| out-width: 80%
#| echo: false
#| message: false
#| warning: false
expand_grid(lv_std = scale(seq(0, 1, 0.1),
                           center = mean(syllables_data$lv),
                           scale = sd(syllables_data$lv))[, 1],
            n_syll_std = scale(seq(1, 4, 1),
                               center = mean(syllables_data$n_syll),
                               scale = sd(syllables_data$n_syll))[, 1]) |> 
    add_epred_draws(model_fit_syllables, ndraws = 50,
                    re_formula = NA) |> 
    mutate(lv_std = rescale_variable(lv_std,
                                     mean = mean(syllables_data$lv),
                                     sd = sd(syllables_data$lv)),
           n_syll_std = rescale_variable(n_syll_std,
                                         mean = mean(syllables_data$n_syll),
                                         sd = sd(syllables_data$n_syll)),
           n_syll_std = as.factor(n_syll_std)) |> 
    ggplot(aes(lv_std, .epred, colour = n_syll_std)) +
    geom_point(data = filter(syllables_data, n_syll < 5),
               aes(x = lv,
                   y = freq_syll,
                   colour = as.factor(n_syll)),
               shape = 1,
               stroke = 1,
               alpha = 1/2) +
    geom_line(aes(group = interaction(.draw, n_syll_std)),
              alpha = 1/10) +
    labs(x = "Cognateness (Levenshtein similarity)",
         y = "Syllabic frequency",
         colour = "Number of syllables") +
    scale_color_manual(values = clrs) +
    scale_y_continuous(n.breaks = 8) +
    theme(panel.grid = element_blank())
```

>>>>

items <- items |> 
    mutate(xsampa = ipa::ipa(ipa)) |> 
    mutate(syll = strsplit(xsampa, 
                           split = paste0(.sep = c("\\.", "\\\""), 
                                          collapse = "|")),
           syll = lapply(syll, function(x) x[x != ""]),
           n_syll = map_int(syll, length),
           item = str_remove(item, "cat_|spa_"))
```

@@

```{r tbl-items-table}
#| label: tbl-syll-items
#| tbl-cap: Sample of items included in the BVQ questionnaire and their syllabified SAMPA transcriptions in Catalan and Spanish
#| echo: false
#| warning: false
#| message: false
items |> 
    distinct(te, item, language, meaning, xsampa, n_syll) |> 
    left_join(distinct(pool_tmp, te, language, item, label),
              by = join_by(te, item, language)) |> 
    pivot_wider(names_from = language,
                values_from = c(item, label, xsampa, n_syll),
                id_cols = c(te, meaning),
                names_repair = janitor::make_clean_names,
                values_fn = first) |> 
    sample_n(15) |> 
    gt() |> 
    cols_hide(c(te, item_catalan, item_spanish)) |> 
    tab_spanner("Catalan", matches("catalan")) |> 
    tab_spanner("Spanish", matches("spanish")) |> 
    cols_label(meaning = "Translation",
               label_catalan = "Item",
               xsampa_catalan = "X-SAMPA",
               n_syll_catalan = "Syllables",
               label_spanish = "Item",
               xsampa_spanish = "X-SAMPA",
               n_syll_spanish = "Syllables") |> 
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners()) |> 
    tab_style(cell_text(style = "italic"),
              cells_column_labels())
```

@@

```{r fig-syll-number}
#| label: fig-syll-number
#| fig-cap: Distribution of the number of syllables in Catalan and Spanish
#| fig-width: 7
#| fig-height: 3
#| echo: false
#| warning: false
#| message: false
pool_tmp |> 
    group_by(language) |> 
    add_count(language, name = "n_lang") |> 
    ungroup() |> 
    count(language, n_lang, n_syll) |> 
    mutate(prop = n/n_lang) |> 
    ggplot(aes(n_syll, n)) +
    facet_wrap(~ language) +
    geom_col(colour = "white", fill = "grey70") +
    geom_text(aes(label = scales::percent(prop)),
              size = 3,
              position = position_nudge(y = 15),
              colour = "black") +
    labs(x = "Number of syllables",
         y = "Number of words",
         fill = "Language") +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())
```

@@

```{r fig-top-syll-freq}
#| label: fig-top-syll-freq
#| fig-cap: "Top 20 most frequent syllables"
#| fig-width: 8
#| fig-height: 3
#| out-width: 80%
#| echo: false
#| warning: false
#| message: false
syll |> 
    group_by(language) |> 
    top_n(20, freq_syll) |> 
    ggplot(aes(tidytext::reorder_within(syll, desc(freq_syll), language),
               freq_syll)) +
    facet_wrap(~ language,
               scales = "free_x",
               ncol = 1) +
    geom_col(colour = "white",
             fill = "grey70") +
    labs(x = "Syllable",
         y = "Relative frequency\n(counts per million)",
         colour = "Language",
         fill = "Language") +
    tidytext::scale_x_reordered() +
    scale_y_continuous(labels = \(x) format(x, big.mark = ",")) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 10),
          panel.grid = element_blank())
```

@@

```{r syllables-posterior}
#| label: syllables-posterior
#| echo: false
#| warning: false
#| message: false
rope_interval <- c(lower = -0.5, upper = 0.5)
# tidy predictor names
str_repl <- c(
    "b_Intercept" = "Intercept",
    "b_n_syll_std" = glue::glue("Syllables (+1 SD, {round(sd(syllables_data$n_syll), 3)})"),
    "b_lv_std" = glue::glue("Cognateness (+1 SD, {round(sd(syllables_data$lv), 2)})")
)

post_syllables <- model_fit_syllables |>
    bayestestR::describe_posterior(test = "rope",
                                    rope_range = rope_interval,
                                    ci_method = "HDI") |>
    as_tibble() |>
    janitor::clean_names() |>
    mutate(.variable_name = factor(
        parameter,
        levels = names(str_repl),
        labels = str_repl) |>
            as.character(),
        parameter = ifelse(grepl("Intercept", parameter),
                           stringr::str_remove_all(parameter, "Intercept \\(|\\)"),
                           parameter)) |>
    select(.variable = parameter,
           .variable_name,
           .median = median,
           .upper = ci_high,
           .lower = ci_low,
           .rope_overlap = rope_percentage) |>
    ungroup()

@@

@fig-syll-marginal shows the median posterior-predicted syllabic frequencies for words with one to four syllables, for the whole range of cognateness values. Overall, cognate words' syllabic frequency is equivalent to that of non-cognates. This suggests that the cognate facilitation effect in word acquisition reported in the present study is not the result from an association between cognateness and higher syllabic frequencies.

```{r fig-syll-marginal}
#| label: fig-syll-marginal
#| fig-cap: "Posterior-predictions of the syllabic frequency model. Thicker lines indicate the median of the posterior predictions, and thinner lines indicate individual posterior predictions."
#| fig-width: 7
#| fig-height: 5
#| out-width: 80%
#| echo: false
#| message: false
#| warning: false
expand_grid(lv_std = scale(seq(0, 1, 0.1),
                           center = mean(syllables_data$lv),
                           scale = sd(syllables_data$lv))[, 1],
            n_syll_std = scale(seq(1, 4, 1),
                               center = mean(syllables_data$n_syll),
                               scale = sd(syllables_data$n_syll))[, 1]) |> 
    add_epred_draws(model_fit_syllables, ndraws = 50,
                    re_formula = NA) |> 
    mutate(
        lv_std = (lv_std * sd(syllables_data$lv)) + mean(syllables_data$lv),
        n_syll_std = (n_syll_std * sd(syllables_data$n_syll)) + mean(syllables_data$n_syll),
        n_syll_std = as.factor(n_syll_std)) |> 
    ggplot(aes(lv_std, .epred, 
               colour = n_syll_std)) +
    geom_point(data = filter(syllables_data, n_syll < 5),
               aes(x = lv,
                   y = freq_syll,
                   colour = as.factor(n_syll),
                   shape = as.factor(n_syll)),
               alpha = 1/2) + 
    geom_line(aes(group = interaction(.draw, n_syll_std)),
              alpha = 1/10) +
    labs(x = "Cognateness (Levenshtein similarity)",
         y = "Syllabic frequency",
         shape = "Number of syllables",
         colour = "Number of syllables") +
    scale_colour_grey(start = 0, end = 0.7) +
    scale_y_continuous(n.breaks = 8) +
    theme(panel.grid = element_blank(),
          legend.position = "top")
```

### 2023-05-24 manuscript/appendix.qmd
Add new corrections to discussion and reorganise project structure

---
csl: "../assets/apa7.csl"
bibliography: "../assets/references.bib"
link-citations: true
editor: source

format:
  pdf:
    documentclass: article
    papersize: letter
    fontsize: 12pt
    linestretch: 1.5
    date-format: full
    template-partials:
    - partials/title.tex
    - partials/before-body.tex
    include-in-header: include-in-header.tex
    header-includes: \usepackage{tipa}
    geometry:
    - top=1in
    - bottom=1in
    - left=1in
    - right=1in
    colorlinks: true
    linkcolor: DarkSlateBlue
    urlcolor: DarkSlateBlue
    citecolor: DarkSlateBlue
    link-citations: true
    biblio-title: References
    biblio-config: false
    cite-method: biblatex
    biblatex-chicago: false
    biblio-style: apa
    biblatexoptions: backend=biber
    block-headings: false
    indent: true
---

@@

library(targets)

tar_config_set(store = here::here('_targets'),
               script = here::here('_targets.R'))

tar_load_globals()

tar_load(c(bvq_data, childes, items, participants, responses))
tar_load(c(starts_with("model_"), model_loos))
tar_load(c(posterior_draws, marginal_effects_epreds))

@@

clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

color_scheme_set(rep(clrs[1], 6))

my_theme <- theme_minimal() +
    theme(panel.grid = element_line(colour = "grey",
                                    linetype = "dotted"),
          axis.line = element_line(colour = "black"),
          text = element_text(size = 12, colour = "black"),
          axis.text = element_text(colour = "black"))

theme_set(my_theme)

bayesplot_theme_set(my_theme)

@@

options(ggplot2.ordinal.fill = clrs,
        ggplot2.ordinal.colour = clrs,
        ggplot2.discrete.fill = clrs,
        ggplot2.discrete.colour = clrs,
        ggplot2.continuous.fill = scale_color_gradient,
        ggplot2.continuous.colour = scale_color_gradient,
        knitr.graphics.error = FALSE) 
```

# Appendix A: model predictors {.appendix .unnumbered}

@@

# Appendix B: model details {.appendix .unnumbered}

@@

# Appendix C: convergence diagnostics {.appendix .unnumbered}

@@

# Appendix D: frequency and language exposure as separate predictors {.appendix .unnumbered}

@@

# Appendix E: frequency and language exposure as separate predictors {.appendix .unnumbered}

@@

>>>>

---
title: "Appendix"
csl: "../assets/apa7.csl"
bibliography: "../assets/references.bib"
link-citations: true
editor: source

fig-align: center
fig-dpi: 800
echo: false
warning: false
message: false

format:
  pdf:
    documentclass: article
    papersize: letter
    fontsize: 12pt
    linestretch: 1.5
    header-includes:
    - \usepackage{tipa}
    geometry:
    - top=1in
    - bottom=1in
    - left=1in
    - right=1in
    colorlinks: true
    linkcolor: DodgerBlue
    urlcolor: DodgerBlue
    citecolor: DodgerBlue
    link-citations: true
    biblio-config: false
    block-headings: false
    indent: true
---

@@

library(patchwork)
library(gt)
library(ggplot2)

targets::tar_config_set(store = here::here("_targets"),
                        script = here::here("_targets.R"))

targets::tar_load_globals()

targets::tar_load(c(bvq_data, childes, items, 
                    participants, responses))

targets::tar_load(c(model_fit, model_summary, model_epreds,
                    model_convergence, model_fit_prior, 
                    model_ppcs, posterior_doe_summary,
                    posterior_doe_draws))

targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables))

@@

my_theme <- function() {
    theme_ggdist() +
        theme(panel.grid = element_line(colour = "grey",
                                        linetype = "dotted"),
              axis.line = element_line(colour = "black"),
              text = element_text(size = 12, colour = "black"),
              axis.text = element_text(colour = "black"))
}

theme_set(my_theme())

bayesplot::bayesplot_theme_set(my_theme + theme(panel.grid = element_blank()))

@@

options(knitr.graphics.error = FALSE)
```

## Appendix A: model predictors {.appendix .unnumbered}

@@

## Appendix B: model details {.appendix .unnumbered}

@@

## Appendix C: convergence diagnostics {.appendix .unnumbered}

@@

## Appendix D: frequency and language exposure as separate predictors {.appendix .unnumbered}

@@

## Appendix E: frequency and language exposure as separate predictors {.appendix .unnumbered}

@@

## References
